---
title: 'Data From: TERRA REF, An Open Reference Data Set From High Resolution Genomics,
  Phenomics, and Imaging Sensors'
author: "David LeBauer, long list TBD"
date: "2/10/2020"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = FALSE, message = FALSE)
```


```{r eml-metadata}
library(EML)
dsl <- eml$creator(individualName = eml$individualName(givenName = "David",
                                                       surName = "LeBauer"),
                   organizationName = "University of Arizona",
                   electronicMailAddress = "dlebauer@arizona.edu",
                   userId = list(directory = "https://orcid.org",
                                 userId    = "https://orcid.org/0000-0001-7228-053X"))
                   
doc <- list(packageId = "dataset-1", system = "Dryad",
            dataset = eml$dataset(
              title = "Data From: TERRA REF, An Open Reference Data Set From High Resolution Genomics, Phenomics, and Imaging Sensors",
              creator = dsl,
              contact = dsl))

```

## Abstract

```{r abstract, results = 'asis'}
abstract <- EML::set_TextType("./abstract.md")
class(abstract) <- 'abstract'
doc$dataset$abstract <- abstract
writeLines(readLines('abstract.md'))
```

## Coverage

```{r location}
library(traits)
library(dplyr)
library(stringr)
# library(kableExtra)
# library(jsonlite)


options(betydb_url = "https://terraref.ncsa.illinois.edu/bety/",
        betydb_api_version = 'v1',
        betydb_key = readLines('~/.betykey', warn = FALSE))

mac_field <- betydb_query(table = 'sites', sitename = '~MAC Field Scanner Field')
library(sf)

mac_bounds <- st_bbox(st_as_sfc(mac_field$geometry))

sorghum <- betydb_query(table = 'species', genus = 'Sorghum')

species <-   data.frame(Genus = sorghum$genus, 
                          Species = sorghum$scientificname)

doc$dataset$coverage <- set_coverage(beginDate = "2017-04-13",
                                     endDate = "2018-08-02",
                                     geographicDescription   = "Maricopa Agricultural Center, Maricopa, AZ",
                                     sci_names = species,
                                     westBoundingCoordinate  = mac_bounds['xmin'],
                                     eastBoundingCoordinate  = mac_bounds['xmax'],
                                     northBoundingCoordinate = mac_bounds['ymax'],
                                     southBoundingCoordinate = mac_bounds['ymin']
                                     #include each plot here?
                                     )
                                     
                                     
```

## Methods

### Experimental Design

```{r experimental-design, results = 'asis'}

experiment_descriptions <- readLines('metadata/experiment_descriptions.txt')

writeLines(gsub("##", "\n####", experiment_descriptions))
```

```{r methods-eml}
doc$dataset$methods <- EML::set_methods(
  sampling_file = 'metadata/experiments_descriptions.txt',
  methods_file = 'metadata/methods.csv')
```

### Cultivars

A list of cultivars and the experiments in which they are planted are listed in the file 'germplasm.csv'. Additional attributes are listed in the files 'season_4_germplasm.json' and 'season_6_germplasm.json'.  

```{r cultivars, eval=FALSE}
readr::read_csv('metadata/germplasm.csv') %>% 
  knitr::kable()
```


```{r cultivars-eml}
# atts_shiny <- EML::shiny_attributes(readr::read_csv('data/germplasm.csv'))

doc$dataset$dataTable <- eml$dataTable(
  entityName = 'metadata/germplasm.csv',
  entityDescription = paste0("List of Sorghum genotypes ('accessions')",
                             "and experiments in which they were included"))

#atts <- read.csv('metadata/germplasm_Attributes_Table.csv')
#doc$dataset$dataTable$attributeList <- set_attributes(atts)
```

```{r}
eml_validate(doc)
write_eml(doc, "metadata/eml_metadata.xml")
```

## Sensor Derived Phenotypes

The following records were generated from sensors by algorithms:

```{r sensor-data}
trait_counts <- readr::read_csv('metadata/trait_counts.csv')

sensor_methods <- c("Green Canopy Cover Estimation from Field Scanner RGB images", 
"3D scanner to 98th quantile height", "Scanner 3d ply data to height", 
"Stereo RGB data to emergence count", "3D scanner to leaf angle distribution", 
"3D scanner to leaf length and width", "3D scanner to panicle count faster_rcnn + roughness threshold + convex hull", "Mean temperature from infrared images")

trait_counts %>% 
  filter(method_name %in% sensor_methods) %>% 
  arrange(desc(n)) %>% 
  knitr::kable()
```
```{r manual-data-counts}
trait_counts %>% 
  filter(!method_name %in% sensor_methods) %>%
  arrange(desc(n)) %>% 
  knitr::kable()
```

## Usage Notes

1. Technical Documentation
  * website: https://docs.terraref.org/
  * source: github.com/terraef/documentation
  * archive: David LeBauer, Craig Willis, Rachel Shekar, Max Burnette, Ting Li, Scott Rohde, et al. 2020. “Terraref/documentation: Season 6 Data Publication (2019)”. Zenodo. doi:10.5281/zenodo.597425
2. Tutorials: Tutorials for accessing and understanding data
  * website: https://terraref.org/tutorials
  * source: github.com/terraref/tutorials
  * archive: TBD
  * YouTube: https://www.youtube.com/channel/UComeQAqYR5aZrXN_3K5iFGw

