# Phenotypes 

## Summary of Available Phenotypes

### Plot level phenotype data 

*files* 
* data/season_4_traits.csv.zip 
* data/season_6_traits.csv.zip

*format* csv with the following fields:

* plot 
* lat
* lon
* scientificname
* genotype
* season
* treatment
* date
* time
* year
* month
* trait
* method
* mean
* checked: has this record been checked? 0 = No, 1 = Yes, 
* notes
* author
* method_type


```{r load-phenotypes}
library(dplyr)
library(readr)
col_types <- readr::cols(plot = col_character(),
                         lat = col_double(),
                         lon = col_double(),
                         scientificname = col_character(),
                         genotype = col_character(),
                         season = col_integer(),
                         treatment = col_character(),
                         date = col_date(),
                         time = col_time(),
                         year = col_integer(),
                         month = col_integer(),
                         trait = col_character(),
                         method = col_character(),
                         mean = col_double(),
                         checked = col_double(),
                         notes = col_character(),
                         author = col_character(),
                         method_type = col_character()
)
s4 <-readr::read_csv('../data/season_4_traits.csv', 
                     col_types = col_types
                     )
s6 <-readr::read_csv('../data/season_6_traits.csv')


```

## Sensor Derived Phenotypes

The following records were generated from sensors by algorithms:

```{r sensor-data}
trait_counts <- readr::read_csv('../metadata/trait_counts.csv')

sensor_methods <- c("Green Canopy Cover Estimation from Field Scanner RGB images", 
"3D scanner to 98th quantile height", "Scanner 3d ply data to height", 
"Stereo RGB data to emergence count", "3D scanner to leaf angle distribution", 
"3D scanner to leaf length and width", "3D scanner to panicle count faster_rcnn + roughness threshold + convex hull", "Mean temperature from infrared images")

trait_counts %>% 
  filter(method %in% sensor_methods) %>% 
  arrange(desc(n)) %>% 
  knitr::kable()
```


```{r manual-data-counts}
trait_counts %>% 
  filter(!method %in% sensor_methods) %>%
  arrange(desc(n)) %>% 
  knitr::kable()
```

